---
title: "Packages"
description: "Explore the OpenRBQM ecosystem of open-source R packages for RBQM."
---

```{r setup}
#| echo: FALSE
#| include: FALSE

# Reusable logic lives in R/ (see R/github.R, R/packages.R).
invisible(lapply(list.files("R", pattern = "\\.R$", full.names = TRUE), source))
```

# Overview

The OpenRBQM framework is designed to provide a robust and flexible solution for risk-based quality management. It includes a variety of packages that help streamline processes and ensure compliance with industry standards.

# The {gsm} Ecosystem

The `{gsm.core}` package is a successor to `{gsm}`, which was **deprecated in March 2025**. The contents of `{gsm}` were split into a family of focused packages, each owning one part of the risk-based monitoring workflow.

## Package lineage

```{mermaid}
flowchart TD
    gsm["<b>gsm</b><br/><i>deprecated Mar 2025</i>"]:::deprecated
    gsm --> core["<b>gsm.core</b><br/>analytics engine"]
    gsm --> workr["<b>workr</b><br/>workflow runtime"]
    gsm --> mapping["<b>gsm.mapping</b><br/>data transformation"]
    gsm --> kri["<b>gsm.kri</b><br/>metrics &amp; visualization"]
    gsm --> reporting["<b>gsm.reporting</b><br/>reporting data model"]

    classDef deprecated fill:#dc3545,stroke:#a71d2a,color:#fff;
```

## How the packages fit together

Together the successor packages form a pipeline that turns raw clinical data into monitoring reports. `{workr}` executes the YAML-defined workflows that drive each step, and `{gsm.core}` provides the analytics used to construct and evaluate the metrics.

```{mermaid}
flowchart LR
    raw[("Raw / source<br/>datasets")] --> mapping["gsm.mapping<br/><i>map to domains</i>"]
    mapping --> kri["gsm.kri<br/><i>generate & visualize metrics</i>"]
    kri --> reporting["gsm.reporting<br/><i>reporting data model</i>"]
    reporting --> reports[["Reports"]]

    core["gsm.core<br/><i>analytics engine</i>"] -. "metric analytics" .-> kri
    workr["workr<br/><i>workflow runtime</i>"] -. "executes workflows" .-> mapping
    workr -.-> kri
    workr -.-> reporting

    classDef runtime fill:#0d6efd,stroke:#0a58ca,color:#fff;
    class workr,core runtime;
```

## Package roles

| Package | Role |
|---|---|
| **`{gsm.core}`** | Analytics functionality used to construct and evaluate metrics. |
| **`{workr}`** | Workflow runtime that executes YAML-defined steps and workflows. |
| **`{gsm.mapping}`** | Workflows that transform raw/source datasets into the appropriate domains. |
| **`{gsm.kri}`** | Workflows to generate metrics, plus functionality to visualize and report on them. |
| **`{gsm.reporting}`** | Workflows to build the reporting data model needed to generate reports. |

# Packages
```{r get_packages}
#| echo: false
#| results: asis
#| warning: false

# orgs, non_package_repos, and the curated package vectors come from R/config.R
# (sourced above) so the Packages and Vignettes tabs stay in sync.
packages <- get_package_table(orgs, non_package_repos)

cat(render_section(packages, "Core Packages", core_pkgs))
cat(render_section(packages, "Extensions and Data", ext_pkgs))
cat(render_section(packages, "Apps and Plugins", apps_pkgs))
cat(render_section(packages, "Quality", qual_pkgs))
cat(render_section(packages, "Additional Packages",
                   remaining_packages(packages, c(core_pkgs, ext_pkgs, apps_pkgs, qual_pkgs, deprecated_pkgs))))
cat(render_section(packages, "Deprecated", deprecated_pkgs))
```