---
title: "Packages"
description: "Explore the OpenRBQM ecosystem of open-source R packages for RBQM."
---

```{r load_packages}
#| echo: FALSE
#| include: FALSE

# Load necessary libraries
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(glue)

# Authenticated GET against the GitHub API.
# Uses GITHUB_PAT as a bearer token when set (higher rate limits); otherwise
# falls back to unauthenticated requests so local renders still work.
gh_get <- function(url) {
  pat <- Sys.getenv("GITHUB_PAT")
  if (nzchar(pat)) {
    GET(url, add_headers(Authorization = paste("Bearer", pat)))
  } else {
    GET(url)
  }
}
```

# Overview

The OpenRBQM framework is designed to provide a robust and flexible solution for risk-based quality management. It includes a variety of packages that help streamline processes and ensure compliance with industry standards.

# The {gsm} Ecosystem

The `{gsm.core}` package is a successor to `{gsm}`, which was **deprecated in March 2025**. The contents of `{gsm}` were split into a family of focused packages, each owning one part of the risk-based monitoring workflow.

## Package lineage

```{mermaid}
flowchart TD
    gsm["<b>gsm</b><br/><i>deprecated Mar 2025</i>"]:::deprecated
    gsm --> core["<b>gsm.core</b><br/>analytics engine"]
    gsm --> workr["<b>workr</b><br/>workflow runtime"]
    gsm --> mapping["<b>gsm.mapping</b><br/>data transformation"]
    gsm --> kri["<b>gsm.kri</b><br/>metrics &amp; visualization"]
    gsm --> reporting["<b>gsm.reporting</b><br/>reporting data model"]

    classDef deprecated fill:#dc3545,stroke:#a71d2a,color:#fff;
```

## How the packages fit together

Together the successor packages form a pipeline that turns raw clinical data into monitoring reports. `{workr}` executes the YAML-defined workflows that drive each step, and `{gsm.core}` provides the analytics used to construct and evaluate the metrics.

```{mermaid}
flowchart LR
    raw[("Raw / source<br/>datasets")] --> mapping["gsm.mapping<br/><i>map to domains</i>"]
    mapping --> kri["gsm.kri<br/><i>generate & visualize metrics</i>"]
    kri --> reporting["gsm.reporting<br/><i>reporting data model</i>"]
    reporting --> reports[["Reports"]]

    core["gsm.core<br/><i>analytics engine</i>"] -. "metric analytics" .-> kri
    workr["workr<br/><i>workflow runtime</i>"] -. "executes workflows" .-> mapping
    workr -.-> kri
    workr -.-> reporting

    classDef runtime fill:#0d6efd,stroke:#0a58ca,color:#fff;
    class workr,core runtime;
```

## Package roles

| Package | Role |
|---|---|
| **`{gsm.core}`** | Analytics functionality used to construct and evaluate metrics. |
| **`{workr}`** | Workflow runtime that executes YAML-defined steps and workflows. |
| **`{gsm.mapping}`** | Workflows that transform raw/source datasets into the appropriate domains. |
| **`{gsm.kri}`** | Workflows to generate metrics, plus functionality to visualize and report on them. |
| **`{gsm.reporting}`** | Workflows to build the reporting data model needed to generate reports. |

# Packages
```{r get_packages}
#| echo: false
#| results: asis
#| warning: false

# Function to get package names from a GitHub repository
get_packages <- function(org) {
    url <- paste0("https://api.github.com/orgs/", org, "/repos?type=public")
    response <- gh_get(url)
    content <- content(response, as = "text")
    json <- fromJSON(content)
    return(json)
}

# Repositories to pull packages from
orgs <- c("openrbqm", "gilead-biostats")

# Get packages from each repository
packages <- orgs %>% map(get_packages) %>% bind_rows
packages$description[is.na(packages$description)] <- "No Description Provided"

# Repositories that are not packages / should not be listed here.
# (Workshop repos are surfaced on the Outreach page instead.)
exclude_pkgs <- c(
  "openrbqm.github.io", "cluster", "openrbqm",
  "openrbqm-workshop", "openrbqm-workshop-eu25"
)
packages <- packages %>% filter(!tolower(name) %in% exclude_pkgs)

# Render a subset of packages as a responsive grid of cards (same style as the
# home page). The link text is the package name only (org prefix removed).
render_cards <- function(df) {
  cat("\n::: {.grid}\n\n")
  df %>%
    glue_data(
      "::: {{.g-col-12 .g-col-md-6 .g-col-lg-4}}\n",
      "::: {{.card .h-100 .nav-card}}\n",
      "::: {{.card-body}}\n",
      "#### [{name}]({html_url})\n\n",
      "{description}\n\n",
      ":::\n:::\n:::\n"
    ) %>%
    cat(sep = "\n")
  cat("\n:::\n\n")
}

# Render a curated section: cards for `pkg_names` (in that order) under `heading`.
render_section <- function(heading, pkg_names) {
  sub <- packages %>%
    filter(tolower(name) %in% pkg_names) %>%
    arrange(match(tolower(name), pkg_names))
  if (nrow(sub) > 0) {
    cat(glue("\n## {heading}\n\n"))
    render_cards(sub)
  }
}

# Curated sections, shown first in the order defined here.
core_pkgs       <- c("gsm.core", "workr", "gsm.mapping", "gsm.kri", "gsm.reporting")
ext_pkgs        <- c("gsm.qtl", "gsm.viz", "clindata", "gsm.datasim")
apps_pkgs       <- c("gsm.app", "gsm.digitpref", "gsm.ae", "gsm.query", "gsm.pd")
qual_pkgs       <- c("qcthat", "gsm.qc", "gsm.utils")
deprecated_pkgs <- c("gsm")

render_section("Core Packages", core_pkgs)
render_section("Extensions and Data", ext_pkgs)
render_section("Apps and Plugins", apps_pkgs)
render_section("Quality", qual_pkgs)

# All remaining (non-curated, non-deprecated) packages.
curated <- c(core_pkgs, ext_pkgs, apps_pkgs, qual_pkgs, deprecated_pkgs)
rest <- packages %>% filter(!tolower(name) %in% curated)

if (nrow(rest) > 0) {
  cat("\n## Additional Packages\n\n")
  render_cards(rest)
}

# Deprecated packages, shown last.
render_section("Deprecated", deprecated_pkgs)

```

# Package Documentation

```{r define article functions, echo=FALSE, results = "asis"}
get_articles <- function(org, repo) {
    url <- paste0("https://api.github.com/repos/",org,"/",repo,"/contents/vignettes")
    response <- gh_get(url)
    if (response$status_code == 200) {
    content <- content(response, as = "text")
    json <- fromJSON(content) %>%
      dplyr::filter(grepl("Rmd", name)) %>%
      dplyr::mutate(name_html = gsub("Rmd", "html", name))
    return(json)
    }
    else{
      return("No articles to display for this repository\n\n")
    }
}

get_meta <- function(article, field) {
  field_line <- article[grep(paste0(field,": "), article)[1]]
  if(length(field_line) > 0) {
    clean_field <- gsub(paste0(field,": "), "", field_line)
    clean_field <- gsub('\"', "", clean_field)
  } else {
    clean_field <- NA
  }
  return(clean_field)
}

get_pkgdown_url <- function(org, repo) {
  url <- paste0("https://api.github.com/repos/",org,"/",repo)
    response <- gh_get(url)
    content <- content(response, as = "text")
    json <- fromJSON(content)
    return(json$homepage)
}
```



```{r get_articles, echo=FALSE, results = "asis"}
packages_clean <- packages %>%
  filter(grepl("/", full_name)) %>%  # remove malformed full_name entries
  mutate(
    org = map_chr(full_name, ~ strsplit(.x, "/")[[1]][1]),
    repo = map_chr(full_name, ~ strsplit(.x, "/")[[1]][2])
  ) %>% 
  select(full_name, org, repo)

# Iterate using pwalk
pwalk(packages_clean, function(full_name, org, repo, ...) {
  # Print header with appropriate spacing
  homepage_link <- get_pkgdown_url(org, repo)
  cat(glue("\n\n## {full_name} \n\n [{repo} Website]({homepage_link})\n\n"))
  cat("\n\n")
  
  articles_meta <- get_articles(org, repo)

  if (is.data.frame(articles_meta)) {
    articles <- map(articles_meta$download_url, readLines)
    names(articles) <- articles_meta$name

    # Add metadata
    articles_meta$title <- map2_chr(articles_meta$name, articles, ~ get_meta(.y, "title"))
    articles_meta$description <- map2_chr(articles_meta$name, articles, ~ get_meta(.y, "description"))
    articles_meta$homepage_link <- homepage_link

    # Derive name_html if needed
    if (!"name_html" %in% colnames(articles_meta)) {
      articles_meta$name_html <- articles_meta$name
    }

    # Output
    glue_data(
      articles_meta,
      '- [{title}]({homepage_link}articles/{name_html}) {description}'
    ) %>% cat(sep = "\n\n")
  } else {
    cat(articles_meta, sep = "\n")
  }
})

```

# Examples

Under Construction